---
title: "Cross Validation"
author: "Alice Tivarovsky"
date: "11/14/2019"
output: github_document
editor_options: 
  chunk_output_type: console
---

# Slides 

Model selection is hard, decide a priori what goes into your model. 

If the model is nested, you can run some tests. If not, there are no tests, but you can use AIC/BIC (balance goodness of fit and complexity). 

Cross Validation: You split full dataset into training dataset and testing dataset (usually an 80/20 split). Training is what you use to build your model. Testing is how you evaluate fit. You evaluate the model using the root mean squared error. 

Cross validation can be used to compare traditional models, in which case it boils down to AIC/BIC. It's very popular in more modern modeling methods (lasso, additive models, regression trees). Here, prediction is the goal. Hypothesis testing is ok but prediction is a lot cooler. 

Tools for cross-validation: 

* lots of helpful functions in modelr
  + add_predictions() and add_residuals()
  + rmse()
  + cross_mc()

# Examples

Setup code

```{r setup}

library(tidyverse)
library(mgcv)
library(modelr)

set.seed(1)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))

```


## Cross Validation

Generate a dataset. Thiis is a non-linear dataset, where a line doesn't quite cut it. 

```{r}
nonlin_df = 
  tibble(
    id = 1:100,
    x = runif(100, 0, 1),
    y = 1 - 10 * (x - .3) ^ 2 + rnorm(100, 0, .3)
  )

nonlin_df %>% 
  ggplot(aes(x = x, y = y)) + 
  geom_point() + theme_bw()
```

Training and testing "by hand"
I’ll split this data into training and test sets (using anti_join!!), and replot showing the split. Our goal will be to use the training data (in black) to build candidate models, and then see how those models predict in the testing data (in red).

```{r}
train_df = sample_frac(nonlin_df, frac = 0.8)
test_df = anti_join(nonlin_df, train_df, by = "id")

ggplot(train_df, aes(x = x, y = y)) + 
  geom_point() + 
  geom_point(data = test_df, color = "red")
```


# Other Materials

- This page touches on ideas that arise in several chapters on modeling in R for Data Science. These tend to assume that this is your first exposure to linear models but good reading:
  - [Intro to modeling](https://r4ds.had.co.nz/model-intro.html)
  - [Basics](https://r4ds.had.co.nz/model-basics.html)
  - [Many models](https://r4ds.had.co.nz/many-models.html)
- The modelr package also has a [website](https://modelr.tidyverse.org/)