---
title: "Linear Models"
author: "Alice Tivarovsky"
date: "11/12/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Slides 

LM = Linear Models
GLM = Generalized Linear Models (include logistic)

# Examples

Setup code

```{r setup}

library(tidyverse)
library(p8105.datasets)

set.seed(1)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))

```


## Model Fitting

The code below loads and cleans the Airbnb data, which we’ll use as a primary example for fitting linear models.

```{r}
data("nyc_airbnb")

nyc_airbnb = 
  nyc_airbnb %>% 
  mutate(stars = review_scores_location / 2) %>% 
  rename(
    boro = neighbourhood_group,
    neighborhood = neighbourhood) %>% 
  filter(boro != "Staten Island") %>% 
  select(price, stars, boro, neighborhood, room_type)

```


Fitting a linear model for price based on stars and boro. Here, reference category is taken as Bronx: 

```{r}
fit = lm(price ~ boro + stars, data = nyc_airbnb)
fit
```

You can run a summary (old-school) and pull out the coefficients: 

```{r}
summary(fit)
coef(fit)
summary(fit)$coef
```

But it's better to use broom::tidy to create a workable dataframe. 

```{r}
fit %>% 
  broom::tidy() %>% 
  mutate(str_replace(term, "boro", "Boro: ")) %>% 
  knitr::kable(digits = 3)
```


